GPU-based Adaptive Beamforming for Active Sonar Imaging

### 2012-05-08 ECUA2012/gpu_capon_ecua2012.tex
la2pdf commit - 2012:05:08	04:55:41	(Tue 8.May)

We have tested our \gls{GPU} implementation of the \gls{MVDR} beamformer on two experimental datasets from the 32 element Kongsberg Maritime HISAS1030 sonar~\cite{Hansen2009}. The sonar was attached to the HUGIN \gls{AUV}, and used in sidescan mode to image the 1500 dwt oil tanker wreck Holmengraa (\Fig{holmengraa}). Holmengraa is roughly 68\,m long and 9\,m wide, and lies on a slanted seabed at 77\,m depth. The \gls{MVDR} image was created with the parameters $L=16$, $N_\text{avg}=3$, and $d=1\%$, which proved to be a reasonable selection for this scenario.

@@

If the HISAS1030 was attached to a platform moving at 1.8m/s, for which the maximum range is 250\,m~\cite{Hansen2010}, the pulse repetition frequency could be set to 3\,Hz. Furthermore, the HISAS1030's bandwidth is 30\,kHz, so we set this to be the complex sampling frequency, and beamform $M=32$ lateral beams. The throughput required for realtime processing of these images would then be 1\,Mpixels/s. As we have seen, this is something our \gls{GPU} implementation could handle, leaving the \gls{CPU} free to take on other assignments.

@@

The \gls{MVDR} beamformer is an algorithm well suited for implementation on a \gls{GPU}. This is because the computations are independent on a pixel level, and partially also within each pixel. We were able to achieve speedup factors of 2-3 orders of magnitude by implementing our active sonar capable \gls{MVDR} implementation on a high-end \gls{GPU}. This performance is sufficient for computing critically sampled full-coverage sectorsscan images from the HISAS1030 sonar in realtime. Furthermore, offloading tasks to the \gls{GPU} might be desirable in itself since sonar processing chains are often \gls{CPU} intensive.

>>>>

We have tested our \gls{GPU} implementation of the \gls{MVDR} beamformer on two different experimental datasets from the 32 element Kongsberg Maritime HISAS1030 sonar~\cite{Hansen2009}. The sonar was attached to the HUGIN \gls{AUV}, and one one of the datasets used in sidescan mode to image the 1500 dwt oil tanker wreck Holmengraa (\Fig{holmengraa}). Holmengraa is roughly 68\,m long and 9\,m wide, and lies on a slanted seabed at 77\,m depth. The \gls{MVDR} image was created with the parameters $L=16$, $N_\text{avg}=3$, and $d=1\%$, which proved to be a reasonable selection for this scenario.

@@

If the HISAS1030 was attached to a platform moving at 1.8m/s, for which the maximum range is 250\,m~\cite{Hansen2010}, the pulse repetition frequency could be set to 3\,Hz. We further assume a critical complex sampling frequency equal to the HISAS1030's bandwidth of 30\,kHz, and beamform $M=32$ lateral beams. The throughput required for realtime processing of these images would then be 1\,Mpixels/s. As we have seen, this is something our \gls{GPU} implementation can handle if $L$ is not too large.%, leaving the \gls{CPU} free to take on other assignments.

@@

The \gls{MVDR} beamformer is an algorithm well suited for implementation on a \gls{GPU}. This is because the computations involved are independent on a pixel level, and partially also within each pixel. We were able to achieve speedup factors of 2-3 orders of magnitude by implementing our active sonar capable \gls{MVDR} implementation on a high-end \gls{GPU}. This performance is sufficient for computing critically sampled full-coverage sectorsscan images from the HISAS1030 sonar in realtime. Furthermore, even if such processing speeds are not required, it might be advantageous to relieve the \gls{CPU} from some of its workload. After all, these chips are designed to cooperate.
